```{r}
library(dplyr)
library(dbplyr)
library(dbplot)
library(DBI)
library(tidyverse)
library(data.table)
```

```{r}
# Access sqlite DB
weather_con <- DBI::dbConnect(RSQLite::SQLite(), "../output/weather_data.sqlite")
```

```{r}
test <- dbSendQuery(conn = weather_con, statement = "SELECT * from weather where random() <= 0.1 LIMIT 1000")
test_sel <- dbFetch(test)
print(test_sel)
print(colnames(test_sel))

foo <- separate(test_sel, "NAME", c("Station_Name", "Country"), sep= ", ")
print(foo)
```

Station Names:
```{r}
names <- dbSendQuery(conn = weather_con, statement = "SELECT DISTINCT(NAME) from weather ORDER BY NAME ASC")
names_sel <- dbFetch(names)
print(names_sel)
```

Number of days:
```{r}
num_days <- 31 + 366 + 31
```

Main Table:
```{r}
weather <- tbl(weather_con, "weather") %>% select(STATION, DATE, NAME, TEMP, MAX, MIN, FRSHTT)
weather

num_total <- weather %>% summarise(n()) %>% collect()
num_total <- num_total$`n()`
  #%>% add_column(separate(), w = 0)
```

Missing Temp Data?
```{r}
num_missing <- weather %>% filter(TEMP == 9999.9 | MAX == 9999.9 | MIN == 9999.9) %>% summarise(n()) %>% collect()
num_missing <- num_missing$`n()`

perc_missing <- num_missing/num_total; perc_missing
```
Missing Names?
```{r}
name_missing <- weather %>% filter(is.na(NAME)) %>% summarise(n()) %>% collect()
name_missing <- name_missing$`n()`

perc_missing <- name_missing/num_total; perc_missing
```


Our percent of missing data is very small, so we can reasonably get rid of it.
```{r}
weather <- weather %>% filter(TEMP != 9999.9 & MAX != 9999.9 & MIN != 9999.9) %>% filter(!is.na(NAME))
```

Now we want to group the observartions by country, but first we need a separate country column


Create new column as country and return averages of temperature and precipitation:

```{r}
weather <- dbSendQuery(conn = weather_con, statement = "SELECT DATE, SUBSTR(NAME, INSTR(NAME, ', ') + 2) as Country_Code, AVG(TEMP) as Average_Temp, COUNT(TEMP) as Station_Count, AVG(MAX) as Average_Max, AVG(MIN) as Average_Min, AVG(PRCP) as Average_Prcp FROM weather WHERE NAME IS NOT NULL AND TEMP <> 9999.9 AND MAX <> 9999.9 AND MIN <> 9999.9 GROUP BY Country_Code, DATE " )

weather_sel <- dbFetch(weather)
head(weather_sel)
```
Now for the US, the state is listed alongside the country, so we need to split those up into separate columns as well
```{r}
weather_sel <- weather_sel %>% separate("Country_Code", c("State_Name", "Country"), sep= " ", fill = "left")
head(weather_sel)

# Double check that it works for US States
weather_sel %>% filter(Country %like% 'US')
```

Replace country codes for Russia and Japan. 
```{r}
weather_sel <- mutate_if(weather_sel, 
                is.character, 
                str_replace_all, pattern = "RS", replacement = "RU")
weather_sel <- mutate_if(weather_sel, 
                is.character, 
                str_replace_all, pattern = "JA", replacement = "JP")
```


Now all we need to do is reinsert this transformed table into our database

```{r}
dbWriteTable(
  weather_con,
  "weather_transformed",
  weather_sel,
  overwrite=TRUE
)
```

And let's test with a query to ensure it was written correctly
```{r}
test <- dbSendQuery(conn = weather_con, statement = "SELECT DISTINCT(DATE) from weather_transformed")
test_sel <- dbFetch(test)
head(test_sel)
```
Hooray!

Include a second table with US states grouped together
```{r}
weather_sel_grouped <- weather_sel %>% select(!State_Name) %>% group_by(DATE, Country) %>% summarise(DATE = DATE, Station_Count = sum(Station_Count), Average_Temp = mean(Average_Temp), Average_Max = mean(Average_Max), Average_Min = mean(Average_Min), Average_Prcp = mean(Average_Prcp)) %>% distinct()
```

Add this as a new table as well
```{r}
dbWriteTable(
  weather_con,
  "weather_transformed_grouped",
  weather_sel_grouped,
  overwrite=TRUE
)
```

Trying to convert Lat/Longs to Country Names
```{r}
latlong <- dbSendQuery(conn = weather_con, statement = "SELECT LATITUDE, LONGITUDE FROM weather GROUP BY LATITUDE, LONGITUDE " )

latlong <- dbFetch(latlong)
head(latlong)
```

```{r}
latlong <- latlong %>% filter(!is.na(LATITUDE) & !is.na(LONGITUDE))
latlong$LATITUDE <- as.numeric(latlong$LATITUDE)
latlong$LONGITUDE <- as.numeric(latlong$LONGITUDE)
head(latlong)

# From https://stackoverflow.com/questions/14334970/convert-latitude-and-longitude-coordinates-to-country-name-in-r

library(sp)
library(rworldmap)

# The single argument to this function, points, is a data.frame in which:
#   - column 1 contains the longitude in degrees
#   - column 2 contains the latitude in degrees
coords2country = function(points)
{  
  countriesSP <- getMap(resolution='high')
  print(countriesSP)

  # convert our list of points to a SpatialPoints object

  #setting CRS directly to that from rworldmap
  pointsSP = SpatialPoints(points, proj4string=CRS(proj4string(countriesSP)))  

  # use 'over' to get indices of the Polygons object containing each point 
  indices = over(pointsSP, countriesSP)

  # return the ADMIN names of each country
  indices$ADMIN  
  #indices$ISO3 # returns the ISO3 code 
  #indices$continent   # returns the continent (6 continent model)
  #indices$REGION   # returns the continent (7 continent model)
}

countries <- coords2country(latlong)
latlong_country <- data.frame(cbind(latlong, countries))
head(latlong_country)
```



